%% acknowledgments.tex

% From mitthesis package
% Version: 1.02, 2024/06/19
% Documentation: https://ctan.org/pkg/mitthesis

\chapter*{Acknowledgments}
\pdfbookmark[0]{Acknowledgments}{acknowledgments}


I would like to express my deepest gratitude to my supervisor, Yoshihiro Nagano, for his excellent guidance throughout this work. His insigthful feedback, support in exploring and refining ideas and the thoughtful comments have been invaluable for me. Especially during times when the results did not align with my expectations he was able to motivate me and always find a new angle to view at the problems that eventually enabled me to make the next step forward. 

I would also like to thank Prof.\ Florentin Wörgötter for taking the role of second supervisor. Even though the topic was new to him, he was willing to support the work without hesitation. This gave me the unique opportunity to explore this interesting field. 

I also want to thank Linnéa Nöth for her support during the entire process. The engaging discussions, her help in structuring both my thoughts and the text, and encouragement during challenging moments kept me going. Thank you!

\section*{Software and Tools Used}

% Für diese Arbeit wurden diverse software und tools genutzt. Darunter die folgenden: Visual Studio Code als Editor für das Manuskript und den Programmcode, GitHub zur Versionierung des Manuskripts und des sourcecodes, Zotero zum Management der Literatur, LaTex zur Erstellung des fertigen Dokuments, ChatGPT zur Optimierung von Programmcode und der Diskussion von Formulierungen in der englischen Sprache, DeepL zur Übersetzung von der deutschen in die englische Sprache. Zusätzlich Google Scholar, Consensus.app und allen.ai paperfinder zur Literaturrecherche. 

Various software and tools were used in the course of this work, including the following: Visual Studio Code as the editor for both the manuscript and source code, GitHub for version control, Zotero for literature management, and LaTeX for compiling the final document. Additionally, ChatGPT was used to optimize programming code and to discuss English phrasings, DeepL supported translations from German into English. For literature research, Google Scholar, Consensus.app, and the Allen.ai PaperFinder were used. The exact libraries that have been used for the programming part of this thesis are listed in the according requirements of the conda environment in the GitHub repository of this work. 
